Light inextensible strings under tension in the Schwarzschild geometry

### 2019-03-20 string_ajp.tex
rearrangement of figures plus typos

Solutions to this differential equation correspond to a taut string
confined to Flamm's paraboloid~\cite{flamm1916}; see
figure~\ref{flamm}.

@@

Taking strings for which $r>1$ first,
figure~\ref{closest_approach_non_self_intersecting} shows a sequence
of non-self-intersecting numerical solutions to
equation~\ref{rdashdash}, rotated so that the tangential point of
closest approach occurs at~$\phi=0$.
Figure~\ref{closest_approach_self_intersecting} shows the
corresponding diagram for self-interesecting strings
(figure~\ref{light_closest_approach} shows comparable null geodesic
diagram for comparison).  A bisection technique reveals that if the
closest approach is greater than a critical value of~$\simeq
1.076109317(1)$, the string does not self-intersect; if the closest
approach is less than this value, then the string crosses itself
at~$\phi=\pi/2$.  As discussed above, if $\left(r'\right)^2=2r(r-1)$,
the string has a point of inflection at which point the radius of
curvature becomes infinite.  Figure~\ref{curvature_switch} shows
non-self intersecting strings rotated so that the point of inflection
occurs at~$\phi=0$; we see that each string has a finite length in
which it is convex to the origin and two semi-infinite stretches in
which the curvature is negative.

The case where the string is tangent to the event horizon at~$\phi=0$,
including a radial string, is shown in
Figure~\ref{fixed_EH_intersection}.

@@

\begin{figure}[p] % Created by "light_closest_approach.R"
\centering
\includegraphics[width=170mm]{light_closest_approach.pdf}
\caption{Null geodesics in the Schwarzschild geometry passing
  through~$(r,0)$, at which point they are tangential.  Note the
  differences between these curves and the taut strings shown
  elsewhere: unlike taug strings, null geodesics have positive
  curvature everywhere, may cross the event horizon inwards, and are
  never tangential to the event horizon}
\label{light_closest_approach}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics{curvature_switch.pdf} % created by "radius_of_curvature_switch.R"
\caption{Strings arranged by point of curvature switch, which occurs
  at~$\phi=0$.  Colour changes from black to red at points of
  infinite~$R$; black indicates convex to the origin and red indicates
  concave regions}
\label{curvature_switch}
\end{figure}

@@

\begin{figure}[p] % created by "one_free_end_r_equals_2.R"
\centering
\includegraphics[width=170mm]{one_free_end_r_equals_2.pdf}
\caption{Strings passing through $(2,0)$ tangential to event horizon}
\label{one_free_end_r_equals_2}
\end{figure}

\begin{figure}[p] % created by "one_free_end_fixed_EH_intersection.R"
\centering
\includegraphics[width=170mm]{one_free_end_fixed_EH_intersection.pdf}
\caption{Strings tangential to event horizon}
\label{fixed_EH_intersection}
\end{figure}

>>>>

Solutions to this differential equation correspond to a taut string
confined to Flamm's paraboloid~\cite{flamm1916}; see
Figure~\ref{flamm}.

@@

Taking strings for which $r>1$ first,
Figure~\ref{closest_approach_non_self_intersecting} shows a sequence
of non-self-intersecting numerical solutions to
equation~\ref{rdashdash}, rotated so that the tangential point of
closest approach occurs at~$\phi=0$.
Figure~\ref{closest_approach_self_intersecting} shows the
corresponding diagram for self-interesecting strings
(Figure~\ref{light_closest_approach} shows comparable null geodesic
diagram for comparison).  A bisection technique reveals that if the
closest approach is greater than a critical value of~$\simeq
1.076109317(1)$, the string does not self-intersect; if the closest
approach is less than this value, then the string crosses itself
at~$\phi=\pi/2$.  As discussed above, if $\left(r'\right)^2=2r(r-1)$,
the string has a point of inflection at which point the radius of
curvature becomes infinite.  Figure~\ref{curvature_switch} shows
non-self intersecting strings rotated so that the point of inflection
occurs at~$\phi=0$; we see that each string has a finite length in
which it is convex to the origin and two semi-infinite stretches in
which the curvature is negative.

The cases where the string is tangent to the event horizon
at~$\phi=0$, including a radial string, are shown in
Figures~\ref{one_free_end_r_equals_2} and~\ref{fixed_EH_intersection}
for strings tangential to or missing the Schwarzschild radius
respectively.  There are other ways to parametrize the strings; Figure
\ref{strings_r_equals_2}, for example, shows strings passing
through~$(2,0)$; Figure~\ref{light_r_equals_2} shows light paths for
comparison.

@@

\begin{figure}[p] % Created by "light_closest_approach.R"
\centering
\includegraphics[width=170mm]{light_closest_approach.pdf}
\caption{Null geodesics in the Schwarzschild geometry, tangential
  at~$\phi=0$.  Note the differences between these curves and the taut
  strings shown elsewhere: unlike taug strings, null geodesics have
  positive curvature everywhere, may cross the event horizon inwards,
  and are never tangential to the event horizon}
\label{light_closest_approach}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics{curvature_switch.pdf} % created by "radius_of_curvature_switch.R"
\caption{Strings arranged by point at which the radius of curvature
  switches, which occurs at~$\phi=0$.  Colour changes from black to
  blue at points of infinite~$R$; black indicates convex to the origin
  and blue indicates concave regions}
\label{curvature_switch}
\end{figure}

\begin{figure}[p] % created by "one_free_end_r_equals_2.R"
\centering
\includegraphics[width=170mm]{one_free_end_r_equals_2.pdf}
\caption{Strings tangential to the event horizon passing through $(2,0)$}
\label{one_free_end_r_equals_2}
\end{figure}

\begin{figure}[p] % created by "one_free_end_fixed_EH_intersection.R"
\centering
\includegraphics[width=170mm]{one_free_end_fixed_EH_intersection.pdf}
\caption{Strings tangential to event horizon at $(1,0)$}
\label{fixed_EH_intersection}
\end{figure}

@@
